---
title: "R Notebook"
output: html_notebook
---
Scrpt from: https://osf.io/bjrhw/?view_only=c1d70b530f074cdbb036eccadfb6dd87

Deposited on Github by DVMB 25th July 2020.
Updated to check Ns for writeup - checked exclusions: total with data on both tasks is 151. Corrections made.



```{r checkpackages}

#Check for required R packages and install if not on system

list.of.packages <- c("nlme")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(nlme)
#library(lmerTest)
library(stringr)
library(reshape2)
library(ggplot2)
require(tidyverse)
```

Load in data and clean if necessary.
```{r loaddata}
#rootdir <- "H:/github/DPhil_Chapter4_fTCD/"
rootdir<-paste0(getwd(),'/')
wordgen <- read.csv(paste0(rootdir,'WordGen_results.csv'), stringsAsFactors=FALSE) # The expressive task
pptt <- read.csv(paste0(rootdir,'PPTT_results.csv'), stringsAsFactors=FALSE)       # The receptive task



demographics <- read.csv(paste0(rootdir,'Chpt4_fTCD_demographics.csv'), stringsAsFactors=FALSE)


#Now we have three data frames that we need to merge into fTCD_dat
#This has been altered by DB to keep all variables, as we want to be able to use them in later analyses
#Note the wordgen and pptt variables are automatically renamed in merge as .x and .y

#strip all IDs to first 3 characters - added by DB to check Ns
wordgen$ID<- substr(wordgen$Filename, start = 1, stop = 3)
pptt$ID <- substr(pptt$Filename,start=1,stop=3)
demographics$ID<- substr(demographics$ID,start=1,stop=3)
#There are 154 with data on both tasks and also on demographics

fTCD_dat <- merge(demographics,wordgen,by="ID") 
fTCD_dat <- merge(fTCD_dat,pptt,by="ID")


# Will just delete those with exclusion in either 

bothexclude<-fTCD_dat$exclusions.x+fTCD_dat$exclusions.y
fTCD_dat_short <- fTCD_dat[bothexclude==0, ] # Removes excluded participants

table(fTCD_dat_short$sex,fTCD_dat_short$hand_self_report)

```

# Normality tests
```{r normtests}
shapiro.test(fTCD_dat_short$LI.x[fTCD_dat_short$hand_self_report==1])
shapiro.test(fTCD_dat_short$LI.x[fTCD_dat_short$hand_self_report==0])
shapiro.test(fTCD_dat_short$LI.y[fTCD_dat_short$hand_self_report==1])
shapiro.test(fTCD_dat_short$LI.y[fTCD_dat_short$hand_self_report==0])
```

# Reliability (added by DB)
```{r reliabilities}

wgr <- cor.test(fTCD_dat_short$odd.x,fTCD_dat_short$even.x)
plot(fTCD_dat_short$odd.x,fTCD_dat_short$even.x) #Word gen
wgp <- cor.test(fTCD_dat_short$odd.y,fTCD_dat_short$even.y)
plot(fTCD_dat_short$odd.y,fTCD_dat_short$even.y) #PPTT

pptt2<-read.csv('pptt_retest_results.csv',stringsAsFactors=F)
wg2<-read.csv('wordgen_retest_results.csv',stringsAsFactors=F)
pptt2$ID <- substr(pptt2$Filename,start=8,stop=10)
wg2$ID <- substr(wg2$Filename,start=8,stop=10)

p2 <- select(pptt2,ID,LI)
w2<- select(wg2,ID,LI)
colnames(p2)[2]<-'LI'
colnames(w2)[2]<-'LI'
fTCD_dat_short$wg2.LI <- NA
fTCD_dat_short$pptt2.LI <- NA

#all files are rank ordered so can use %in%

w<-which(fTCD_dat_short$ID %in% w2$ID)
fTCD_dat_short$wg2.LI[w] <- w2$LI
w<-which(fTCD_dat_short$ID %in% p2$ID)
fTCD_dat_short$pptt2.LI[w] <- p2$LI

wgr2 <- cor.test(fTCD_dat_short$LI.x,fTCD_dat_short$wg2.LI)
ppttr2 <- cor.test(fTCD_dat_short$LI.y,fTCD_dat_short$pptt2.LI)

plot(fTCD_dat_short$LI.x,fTCD_dat_short$wg2.LI)
plot(fTCD_dat_short$LI.y,fTCD_dat_short$pptt2.LI)

```

#Preregistered Hypotheses
see https://osf.io/bcxus?view_only=c1d70b530f074cdbb036eccadfb6dd87

Hypoethesis 1
Stronger left lateralization for the expressive than the receptive task

We fit the heterogenous and homogeneous models, chose the optimal fitting model and then interpret the main effect of task.

Hypothesis 2: stronger left lateralization for right handers than left handers

  Model script to test Hypotheses 1 to 4
#
   Created 26/11/2018.
   Modified 27/11/2018.
   Syntax error fixed 11/12/2018

Following the example procedure described here: https://quantdev.ssri.psu.edu/sites/qdev/files/ILD_Ch06_2017_MLMwithHeterogeneousVariance.html

WE have chosen to use nlme rather than lme4 as the specification of heterogeneous within-person error is more convenient in nlme (lme4 is only implemented via a messy solution).

We fit the heterogenous and homogeneous models, chose the optimal fitting model and then interpret the main effect of handedness.



# Homogeneous variance model
```{r homogmodel}
#select relevant data and make long form

fTCD_dat_x <- select(fTCD_dat_short,ID,hand_self_report,LI.x,LI.y)
fTCD_dat_x$ID<-as.factor(fTCD_dat_x$ID)
fTCD_dat_x$hand_self_report<-as.factor(fTCD_dat_x$hand_self_report)  
colnames(fTCD_dat_x)[3:4]<-c('WordGen','PPTT')


fTCD_mod_dat<- melt(fTCD_dat_x) 
colnames(fTCD_mod_dat) <- c('id','hand','task','LI')
```
In chunk below I just try to work through website explanations

```{r model_explanations}
#Model with homogenous variances
# Common models for between and within person variance

#Minimalist model added from: https://quantdev.ssri.psu.edu/sites/qdev/files/ILD_Ch06_2017_MLMwithHeterogeneousVariance.html


mod0.1<-lme(fixed=LI~1,
          random= ~ 1|id,
          data=fTCD_mod_dat, 
          method="REML")

#This just shows you get the same result with this version
# pdSymm() indicates that the random effects are structured as a positive-definite Symmetric matrix

model.01a = lme(fixed = LI ~ 1,  
               random = list(id = pdSymm(form = ~ 1)),
                       data = fTCD_mod_dat,
                   method = 'REML')
summary(model.01a)
VarCorr(model.01a)

# Heterogeneity for between variance and common within person variance
model.01b = lme(fixed = LI ~ 1,  
                random = list(id = pdDiag(form = ~ Handedness)),
                        data = fTCD_mod_dat,
                    method = 'REML')
summary(model.01b)
VarCorr(model.01b)
#I don't understand this - it shows virtually no variance for Handedness

# Heterogeneity for both between and within person variance
model.01c = lme(fixed = LI ~ 1,  
                random = list(id = pdDiag(form = ~ Handedness)),
                weights = varIdent(form = ~ 1 | Handedness),
                        data = fTCD_mod_dat,
                    method = 'REML')
summary(model.01c)
VarCorr(model.01c)
#Nope - this is different but not sure what it is telling me
```


This is the version we used for the paper, which closely follows worked eg on website

```{r model1}

mod0<-lme(fixed=LI~1+hand+task, 
          random=list(id=pdSymm(form=~1)),
          data=fTCD_mod_dat, 
          method="REML")

#to extract the results with pvalue (t-test for marginal significance of each fixed effect with other fixed effects)
summary(mod0)

VarCorr(mod0)

```

#----------------------------------------------------------------------------------#

# Heterogeneous model variance (between-person) and common within person
```{r heteromodel}
mod1<-lme(fixed=LI ~ 1 + hand + task, 
          random=list(id=pdDiag(form= ~ 0 + hand)),
          data=fTCD_mod_dat,
          na.action="na.exclude",
          method="REML")

summary(mod1)

VarCorr(mod1)


#Likelihood ratio test

anova(mod0,mod1)
```

#----------------------------------------------------------------------------------#

We then can check for heterogeneous within-person: except that with 2 groups will have same DF as model 1 - so not feasible?

# Heterogeneous model variance (within-person)
```{r hetero2}
mod2<-lme(fixed=LI ~ 1 + hand + task, 
          random = list(id=pdSymm(form = ~1 )), 
          weights=varIdent(form=~1 | hand), 
          data=fTCD_mod_dat, 
          na.action=na.exclude, 
          method="REML")

summary(mod2)

VarCorr(mod2)

#extract variances for hand and see if they are different.

#residual std deviation
summary(mod2)$sigma

#Residual variance of right hand (dependent on the order of factors)
(summary(mod2)$sigma*1)^2

#Residual variance of left hand
(summary(mod2)$sigma*coef(mod2$modelStruct$varStruct,uncons=FALSE))^2

```
#----------------------------------------------------------------------------------#

#Likelihood ratio test between homogeneous and heterogeneous withi-person

anova(mod0,mod2)

#------------------------------------------------------- ---------------------------#

# Density plots (added by DB)
I'll do these separately for the 2 tasks.
This actually then looks as if the pattern is a bit different, with unimodal distribution for PPTT

```{r densplots}
# Change density plot fill colors by groups
require(ggpubr) #had ggarrange for multiple panel plots
handcode <- c('Left','Right')

fTCD_mod_dat$Handedness <- as.factor(handcode[fTCD_mod_dat$hand])
p1 <- ggplot(fTCD_mod_dat[fTCD_mod_dat$task=='WordGen',], aes(x=LI, fill=Handedness)) +
  geom_density(alpha=0.4)+
  xlim(-3,6)+
 labs(title="Word generation", x ="")
p2 <- ggplot(fTCD_mod_dat[fTCD_mod_dat$task=='PPTT',], aes(x=LI, fill=Handedness)) +
  geom_density(alpha=0.4)+
  xlim(-3,6)+
   labs(title="Semantic Association",
        x ="Laterality index")
ggarrange(p1,p2, ncol = 1, nrow = 2,common.legend=T,legend="bottom")
ggsave(
  "fTCD_densplots.png",
  width = 5, height = 8,
  dpi = 300
)

```

#Hypothesis 3: Significant correlation between expressive and receptive task

#Test with Pearson's correlation

#Hypothesis 4: More variable relationship between expressive and receptive tasks in

#left handers than right handers

#Test by comparing variability of cook's distance between groups using Fligner-Killeen test

#----------------------------------------------------------------------------------#
```{r corrsH3}
#Pearson's correlation between expressive and receptive tasks

H3_results <- cor.test(fTCD_dat_short$LI.x, fTCD_dat_short$LI.y)

#Plot data
fTCD_dat_short$Handedness <- as.factor(fTCD_dat_short$hand_self_report)
levels(fTCD_dat_short$Handedness)<-handcode
print(ggplot(fTCD_dat_short,aes(y=LI.y,x=LI.x,colour=Handedness))+geom_point(size=2)+theme_bw()+
           labs(title="Laterality Indices",x ="Word Generation",y = "Semantic Association")
      + scale_color_manual(values=c("orange1", "royalblue2")))
```

# Do analysis of outliers
Fit a linear model to both handedness groups

```{r linearmodel}
mymod<-lm(LI.y~LI.x,data=fTCD_dat_short)

#Calculate Cook's Distance for each participant
cooks<-cooks.distance(mymod)
fTCD_dat_short$cooks<-cooks

Outlier <- 4*mean(cooks)
fTCD_dat_short$Outlier<-0
fTCD_dat_short$Outlier[fTCD_dat_short$cooks>Outlier]<-1
table(fTCD_dat_short$Outlier)
fTCD_dat_short$Outlier<-as.factor(fTCD_dat_short$Outlier)

#Run the Fligner-Killeen test
H4_p <- fligner.test(fTCD_dat_short$cooks ~ fTCD_dat_short$Handedness)$p.value
```


#plots
```{r doplots}
pcook <- ggplot(fTCD_dat_short, aes(y = LI.y, x = LI.x, colour = Handedness,pch=Outlier)) +
  geom_point(size = 2) +
  theme_bw() +
  labs( x = "Word Generation LI", y = "Semantic Association LI") +
  scale_color_manual(values = c("orange1", "royalblue2"))
ggsave(
  "fTCD_cooks.png",
  width = 5, height = 5,
  dpi = 300
)  

table(fTCD_dat_short$Outlier,fTCD_dat_short$Handedness)
```



# Pirate Plot
library('yarrr')

pirateplot(data=fTCD_mod_dat, LI~task*hand)


